---
topic: "water"
title: "Long-Term Trends in River Flows in B.C."
format: pdf
---       

## Appendix 1

### Volume and Timing of Flow by River Regime

Trends in several flow metrics varied by river regime. There was a clear difference in trends between rain/mixed and snow-dominated regimes.

### Volume metrics

`r average_direction$n[average_direction$direction == "positive" & average_direction$Regime == "Snow-Dominated - Late Peak"]` stations with late peak flows showed Increasing flows. `r average_direction$n[average_direction$direction == "positive" & average_direction$Regime == "Snow-Dominated - Early Peak"]` stations with regular peak flows also showed an increasing trend. In contrast, rain-dominated and mixed-regime rivers had declining average annual flows.

Low summer flows showed a 5% or greater decline in more than half of rain-dominated rivers (`r Low_cat$n[Low_cat$category == "> 5% decrease" & Low_cat$Regime == "Rain-Dominated"]` stations).

```{r include = is_html, eval=is_html, fig.width = 10, fig.height = 8, echo = FALSE, fig.alt="Volume regime"}
volume_regime
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height = 6, fig.alt="Volume regime", echo = FALSE}
volume_regime
```

### Timing metrics

```{r include = is_html, eval=is_html, fig.width = 10, fig.height = 8, echo = FALSE, fig.alt="Timing regime"}
timing_regime
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height = 6, fig.alt="Timing regime", echo = FALSE}
timing_regime
```

Earlier freshet was occurring in many rivers across all regimes. In particular, rivers with spring freshet (`r freshet_direction$n[freshet_direction$direction == "negative" & freshet_direction$Regime == "Snow-Dominated - Early Peak"]` stations) showed an earlier freshet. Snow-dominated rivers also had earlier starts to the low flow period. This included `r low_date_direction$n[low_date_direction$direction == "negative" & low_date_direction$Regime == "Snow-Dominated - Late Peak"]` stations with late peak flows and `r low_date_direction$n[low_date_direction$direction == "negative" & low_date_direction$Regime == "Snow-Dominated - Early Peak"]` stations with regular peak flows.


### Changes in Volume and Timing by Watersheds

### Flow volume

### Average annual flow

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
average_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Average Annual Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

average_by_zone_direction = average_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
annual_bar_plot
```

-   For average annual flow, the vast majority of stations had non-significant trends. In the south-central and -west hydrologic zones, a number of stations had increasing annual flow trends including the Fraser Plateau (`r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Fraser Plateau" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Fraser Plateau"])` stations) and the west and central South Coast Mountains (`r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Western South Coast Mountains" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Western South Coast Mountains"])` and `r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Central South Coast Mountains" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Central South Coast Mountains"])` stations respectively).

## Peak flow

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
peak_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Peak Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

peak_by_zone_direction = peak_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
peak_bar_plot
```

-   There were no clear geographic patterns in trends in peak flow. The only result to highlight is Western Vancouver Island, which has `r peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island" & peak_by_zone_direction$direction == "positive"]` of `r sum(peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations with increasing and `r peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island" & peak_by_zone_direction$direction == "negative"]` of `r sum(peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations with decreasing annual peak flows.

## Low flow
```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
low_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Low Summer Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

low_by_zone_direction = low_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
low_bar_plot
```

-   Trends is summer low flow are most apparent on Vancouver Island. Western Vancouver Islands monitored rivers are almost all (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western Vancouver Island" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations) are decreasing at more than 5% volume per decade. Eastern Vancouver Island also has a large proportion of river low flows in decline (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Eastern Vancouver Island" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Eastern Vancouver Island"])` stations). Other hydrologic zones with decreasing summer low flow volumes to highlight include the Western South and Central Coast Mountains (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western South Coast Mountains" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western South Coast Mountains"])` and `r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Central Coast Mountains" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Central Coast Mountains"])` stations respectively)

# Flow timing

## Date of freshet

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
freshet_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Date of Freshet") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

freshet_by_zone_direction = freshet_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```

```{r fig.width = 10, fig.height=8, echo = FALSE}
freshet_bar_plot
```

-   Timing of freshet has been getting earlier for stations across most hydrologic zones. North-eastern B.C. was the only region that was immune to this effect, apart from the Rocky Mountains which saw earlier freshet at several stations (`r freshet_by_zone_direction$n[freshet_by_zone_direction$Sub_Basin == "Northern Rocky Mountains" & freshet_by_zone_direction$direction == "negative"]` of `r sum(freshet_by_zone_direction$n[freshet_by_zone_direction$Sub_Basin == "Northern Rocky Mountains"])` stations). This pattern was also missing from Vancouver Island, although this is likely due to the majority of rivers not being snow and glacier-fed - the driver behind the freshet phenomena. 

## Start of low flow period
```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
date_low_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Date of Low Summer Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

date_low_by_zone_direction = date_low_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```

```{r fig.width = 10, fig.height=8, echo = FALSE}
date_low_bar_plot
```

Trends in the timing of the start of low flow period are far less clear than that of timing of freshet. Trends indicating earlier timing of low flows were present in some hydrologic zones, most prominent in the Southern Quesnel Highland (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Southern Quesnel Highland" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Southern Quesnel Highland"])` stations), the Fraser Plateau (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Fraser Plateau" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Fraser Plateau"])` stations) and Nechako Plateau (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Nechako Plateau" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Nechako Plateau"])` stations).


## Appendix 2

```{r station_plots, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results = 'asis', fig.height = 7.5, dev = "pdf"}

# individual well plots
# load('../groundwater-levels-indicator/tmp/well_plots.RData')

knitr_latex_char <- function(x) {
  y <- gsub("\\\\", "\\\\textbackslash{}", x) # backslash has to be first!
  y <- gsub("([#$%&_])", "\\\\\\1", y) # Doesn't deal with { or } because of function{}
  y <- gsub("\\^", "\\\\textasciicircum{}", y)
  y <- gsub("~", "\\\\textasciitilde{}", y)
  return(y)
}

# theme_set(base_theme)

  for (station in stations_sf_filtered$STATION_NUMBER) {
    
    stationname <- stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(STATION_NAME)
    
    regimetype = stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(Regime)
    
    reg = stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(Sub_Basin)
    
    cat(paste0('\\subsubsection*{',reg,": ",
               knitr_latex_char(stationname),
               '}'))
    
    cat(paste0('\\subsubsection*{',"Regime Type: ",
               knitr_latex_char(regimetype),
               '}'))

    mapplot <- readPNG(here(paste0("print_ver/out/figs/maps/",stationname,".png")))
    mapplot = ggdraw() +
      draw_image(mapplot)
    
    hydroplot = readPNG(here(paste0("print_ver/out/figs/hgraphs/",station,".png")))  
    hydroplot = ggdraw() +
      draw_image(hydroplot)
    
    volumeplot = readPNG(here(paste0("print_ver/out/figs/volume_plots/",station,".png")))
    volumeplot = ggdraw() +
      draw_image(volumeplot)
    
    timingplot = readPNG(here(paste0("print_ver/out/figs/timing_plots/",station,".png")))  
    timingplot = ggdraw() +
      draw_image(timingplot)    
    
    grid.arrange(mapplot+ theme(plot.margin = unit(c(-1,0.5,-1,0.01),"cm")),
                 hydroplot+ theme(plot.margin = unit(c(-1,0.01,-1,0.01),"cm")),
                 volumeplot + theme(plot.margin = unit(c(0.5,0.5,-1,0.01),"cm")),
                 timingplot + theme(plot.margin = unit(c(0.5,0.5,0,0.01),"cm")),
                 ncol = 2)
    # 
    cat('\\newpage ')
  }
```
