---
topic: "water"
title: "Long-Term Trends in River Flows in B.C."
format: 
    pdf: 
       template: "template.tex"
       keep-tex: FALSE
       latex_engine: xelatex
       gphxpath: "img/"
       mainfont: "Palatino Linotype"
       sansfont: "Calibri"
       
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->
```
```{r setup, echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}

library(dplyr)
library(ggplot2)
library(grid)
library(RColorBrewer)
library(scales)
library(bcgroundwater)
library(envreportutils)
library(gridExtra)
library(xtable)
library(lubridate)
library(forcats)
library(here)
library(magick)
library(png)
library(cowplot)
library(here)
library(sf)
library(stringr)

knitr::opts_chunk$set(cache = FALSE)

theme_set(theme_classic() +
            theme(text = element_text(colour = "black"),
                  axis.line = element_blank(),
                  axis.ticks = element_blank(),
                  panel.grid.major = element_line(colour = "grey85", size = 0.5,
                                                linetype = 1),
                  panel.grid.minor = element_line(colour = "grey90", size = 0.5,
                                                linetype = 1),
                  panel.grid.major.x = element_blank(),
                  panel.spacing = unit(0.6, "lines"),
                  plot.title = element_text(vjust = 2, hjust = 0.5),
                  axis.title = element_text(vjust = 0.1),
                  legend.position = "bottom", legend.title = element_blank(),
                  legend.text = element_text(size = 12),
                  axis.text.x = element_blank(),
                  strip.background = element_blank()))

label.colour <- "black" 
colour.scale <- brewer.pal(3,"Blues")

updateDate <- "February 2024"

# Load datasets
annual_flow_dat = readRDS(here("app/www","annual_flow_dat.rds"))
stations_sf = st_read(here("app/www","stations.gpkg")) 
mk_results_tbl = readRDS(here("print_ver/out/mk_results_tbl.rds"))
mk_results_monthly_tbl = readRDS(here("print_ver/out/mk_results_monthly_tbl.rds"))

stations_sf_filtered = stations_sf %>%
  filter(keep == 1)

# summary table of number of stations in each category
summary_table_cat = mk_results_tbl %>%
  filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(metric, Regime, category) %>%
  summarise(n = n())

summary_table_direction = summary_table_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(metric, Regime,  direction) %>%
  summarise(n = sum(n))

#Subset Average
average_cat = summary_table_cat %>%
  filter(metric == "Average Annual Flow")

average_direction = summary_table_direction %>%
  filter(metric == "Average Annual Flow")

#Subset Peak Flow
peak_cat = summary_table_cat %>%
  filter(metric == "Peak Flow")

peak_direction = summary_table_direction %>%
  filter(metric == "Peak Flow")

#Subset Low Flow
Low_cat = summary_table_cat %>%
  filter(metric == "Low Summer Flow")

Low_direction = summary_table_direction %>%
  filter(metric == "Low Summer Flow")

#Subset Freshet
freshet_cat = summary_table_cat %>%
  filter(metric == "Date of Freshet*")

freshet_direction = summary_table_direction %>%
  filter(metric == "Date of Freshet*")

#Subset Freshet
low_date_cat = summary_table_cat %>%
  filter(metric == "Start of Low Flow Period")

low_date_direction = summary_table_direction %>%
  filter(metric == "Start of Low Flow Period")

# Results by regime
annual_regime = 

# Monthly results

# Subset median flow by month
summary_table_cat_monthly = mk_results_monthly_tbl %>%
  filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(metric, Month, category) %>%
  summarise(n = n())

summary_table_direction_monthly = summary_table_cat_monthly %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(metric, Month, direction) %>%
  summarise(n = sum(n))

# Load figures
load(here("print_ver/out","figures.Rdata"))

# Load map png (perhaps change this to a nicer looking map?)
leaflet_map = image_read(here("print_ver/out/figs","static_leaflet.png"))
```

-   **River water is one of the most important water sources for people living in British Columbian.** Rivers provide freshwater supply and recreational opportunities for British Columbian people. They also offer a variety of habitats that support a wide range of biodiversity. 

-   **Changes in the timing and volume of river flow can affect both human and natural ecosystems.** These changes can affect our ability to predict and manage seasonal water resources. They may disrupt natural systems that rely on typical flow patterns.
    
-   **Low river flows in summer may reduce the amount of water available for uses.They include agriculture, hydroelectric power generation, industry and communities.** Low flows may impede salmon migration and often link to warmer water temperatures. Reduced river levels can decrease water quality, threatening the health of aquatic ecosystems.
    
-   **Extreme high river flows can pose risks to human life and may lead to economic losses.** Flood events are often linked to rainfall and snowmelt. 
    
-   **Long-term trends in the river flow volume varied by flow metric and regime type.** Trends in peak flow volume are stable across the province. But annual flows are increasing for many snow-dominated rivers. Summer low flow volumes are decreasing, particularly in rivers with rain-dominated regimes. 

-   **Freshet is occurring earlier for a large number of rivers across the province.** This change has a less pronounced effect on the start of low flow period. But several snow-dominated rivers are still showing an earlier start to this period.

-   **There has been a shift in seasonal timing of flow events, as seen in monthly flow trends.** Average and summer peak flows are increasing during the winter and spring (November - May). They are declining in the summer (July - September).

<!-- -   **Trends in timing and volume of river flow vary geographically.** Annual average flows increased in the south-west of the province, while there was a strong trend for lower river flows on Vancouver Island. Timing of freshet was consistently earlier across all regions apart from the North-East.  -->

\newpage
    
# River flows in British Columbia

We uses flow metrics to quantify changes in river flow volume and timing of key flow events. There are `r nrow(stations_sf_filtered)` stations across the Province, or `r nrow(stations_sf)` stations when including upstream stations. These stations have tracked river flow for `r min(stations_sf_filtered$N_years)` to `r max(stations_sf_filtered$N_years)` years between `r min(stations_sf_filtered$Min_Year)` and `r max(stations_sf_filtered$Max_Year)`.

## Types of flow regimes

Flow volume and timing of specific flow events vary based on the source of the rivers water. There are many types of river flow regimes^1^. This analysis divides the regimes into four categories:

- **Snow-Dominated - Early Peak**: Snowmelt in the spring and early summer is a major source of flows. This causes a spring freshet and lower volumes in summer. 

- **Snow-Dominated - Late Peak**: Flows are also driven by snow melt. But, due to higher latitude and elevations at the river source, snowmelt occurs later in the year. This leads to a delayed freshet and later summer low flows

- **Rain-Dominated**: Precipitation is the main source of flows. In B.C., these rivers are in the coastal and island regions. Such rivers typically flow high in the winter and low in the summer.

- **Mixed Regime**: Flows comes from both snowmelt and precipitation. This leads to a freshet event and high river flows in the winter.

# Description of river flow metrics

The metrics calculated at each station aim to quantify two aspects of river flow changes:
1) the volume of river flow and 2) the timing of key river flow events.

## Volume metrics

## Average annual flow

Average annual flow (meters^3^/s) is the daily flow averaged for each station and water year^2^ (October – October).

## Peak annual flow

Volume of peak flow (meters^3^/s) is the highest flow identified from the rolling averages with a 3-day window. This metric represents the peak flow for each station for each water year^2^ (October - October).

## Low summer flow

Volume of low flows (meters^3^/s) represents the lowest flow for each low-flow year^3^ at each station. The metric uses the smallest rolling average with a 7-day window across the summer months (May - October).

## Timing metrics

## Date of Freshet

We define Freshet as the flood of a river, often due to snow melt. We estimate the Freshet date as when river discharge reaches 50% of the total for a specified period. The effects of snowmelt on river flow vary by regime. So the method of calculating Freshet timing differs among regimes. For snow-dominated regimes, we use water year to account for early peak. For late peak, we use the calendar year to reflect the delayed Freshet. For mixed regimes, we exclude the rain-dominated winter months (October - February). Instead, we calculate discharge between March and September to estimate Freshet timing. For rain-dominated stations, we do not calculate Freshet, as it does not take place.

## Start of low flow period

We used the long-term Mean Annual Discharge (MAD) to estimate the beginning of low summer flows. We define the start of the low-flow period as the first date when discharge drops to or below 50% of MAD after peak flow. For most rivers, we use 50% MAD, but for rain-dominated rivers, we apply a lower threshold of 20% MAD.

\newpage

# River Stations

The map below shows all `r nrow(stations_sf_filtered)` downstream river stations. The color of the dots represents the flow regime of the rivers where the stations located. Black lines show the boundaries of the six major river basins. They, from top left to bottom right, are Northwest, Liard, Peace, Coastal, Fraser and Columbia. Grey lines outline 23 sub-basins, each containing at least one monitoring station.
```{r overview_map, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.align = 'left', fig.width = 28, fig.asp=0.75}
plot(leaflet_map)
```

# Trend Results

## Exploring water flow trends

The interactive map below offer several options to explore. 

- To begin, go to 'Trend options' tab in the bottom left corner. You can select a basin of interest from the drop-down menu or by clicking it on the map. Choose annual trend or explore the trend for a specific month by selecting ‘monthly’. Next, navigate to the 'Trend to Display' option to select a flow metric of interest from the dropdown menu. You can choose a timespan: 'Three decades (1992–2022)' or 'All available years'. Check the two boxes to explore recently installed or upstream stations.

- After selecting the 'Trend Options', click on a station of interest on the map. This will display see the flow metric plot in the bottom left corner.

- To explore hydrograph information for the selected station, go to the ‘Station Hydrograph’ tab. 

- To reset the view, select “All” or click on the B.C. button in the top left corner to return to all regions.

## Volume metrics

- Of the `r nrow(stations_sf_filtered)` monitored stations, `r sum(average_direction$n[average_direction$direction == "positive"])`
showed an increase in average annual flows. Among them, `r sum(average_cat$n[average_cat$category == "> 5% increase"])` rose by more than 5% per decade. In contrast, `r sum(average_direction$n[average_direction$direction == "negative"])` stations had decreasing average annual flows. The remaining `r sum(average_direction$n[average_direction$direction == "no change"])` stations had no significant trend.

-   Annual peak flows increased at `r sum(peak_direction$n[peak_direction$direction == "positive"])` stations and decreased `r sum(peak_direction$n[peak_direction$direction == "negative"])` stations.The vast majority (`r sum(peak_direction$n[peak_direction$direction == "no change"])`stations) showed no significant change.

-   For low summer flow volume, `r sum(Low_direction$n[Low_direction$direction == "negative"])` showed a decline, `r sum(Low_cat$n[Low_cat$category == "> 5% decrease"])` of them decreased at a rate greater than 5% per decade.

```{r fig.width = 10, fig.height = 8, echo = FALSE}
mk_all
```
* Date of Freshet plot uses a reduced number of stations. This is because the analysis does not include rain-dominated rivers.

## Timing metrics

-   Freshet has been occurring earlier for many locations.Among them, `r sum(freshet_direction$n[freshet_direction$direction == "negative"])` show an earlier trend. Of these, `r sum(freshet_cat$n[freshet_cat$category == "> 2 days earlier"])` stations are
occurring 2 days earlier per decade. `r sum(freshet_direction$n[freshet_direction$direction == "no change"] + freshet_cat$n[freshet_cat$category == "< 1 days change"])` stations
show little change, with less than 1 days change per decade, or no significant trend. No stations show a late Freshet. 

-   Most locations do not show significant trends in the start date of the low flow period.  Among those that do, `r sum(low_date_direction$n[low_date_direction$direction == "negative"])` stations have earlier low flows, while only `r sum(low_date_direction$n[low_date_direction$direction == "positive"])` stations have later low flows..

# Volume and Timing of Flow by River Regime

Trends in several flow metrics varied by river regime. There is a clear difference in trends between rain/mixed and snow-dominated regime. 

# Volume metrics

`r average_direction$n[average_direction$direction == "positive" & average_direction$Regime == "Snow-Dominated - Late Peak"]` stations with late peak flows show Increasing flows. `r average_direction$n[average_direction$direction == "positive" & average_direction$Regime == "Snow-Dominated - Early Peak"]` stations with regular peak flows also show an increasing trend. In contrast, rain-dominated and mixed-regime rivers have declining average annual flows.

Low summer flows show a 5% or greater decline in more than half of rain-dominated rivers (`r Low_cat$n[Low_cat$category == "> 5% decrease" & Low_cat$Regime == "Rain-Dominated"]` stations). 

```{r fig.width = 10, fig.height = 8, echo = FALSE}
volume_regime
```

# Timing metrics

```{r fig.width = 10, fig.height = 8, echo = FALSE}
timing_regime
```

Earlier freshet is occurring in many rivers across all regimes. In particular, rivers with spring freshet (`r freshet_direction$n[freshet_direction$direction == "negative" & freshet_direction$Regime == "Snow-Dominated - Early Peak"]` stations) show an earlier freshet. Snow-dominated rivers also have earlier starts to the low flow period. This include (`r low_date_direction$n[low_date_direction$direction == "negative" & low_date_direction$Regime == "Snow-Dominated - Late Peak"]` stations with late peak flows and `r low_date_direction$n[low_date_direction$direction == "negative" & low_date_direction$Regime == "Snow-Dominated - Early Peak"]` stations with regular peak flows.

# Changes in Flow Volume by Month

## Average monthly flows

```{r fig.width = 10, fig.height=8, echo = FALSE}
monthly_average_bar_plot
```

-   Changes in average river flow across years vary by month. Many rivers show increasing average flows in winter and spring months (December - May). For example, January flows are increasing at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Average" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` stations. In April and May, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Apr" & summary_table_cat_monthly$category == "> 5% increase"]` and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "May" & summary_table_cat_monthly$category == "> 5% increase"]` stations show a more than 5% increase per decade.

## Monthly low flows

```{r fig.width = 10, fig.height=8, echo = FALSE}
monthly_low_flow_bar_plot
```

-   Low flows have increased in January and April at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Apr" & summary_table_direction_monthly$direction == "positive"]` stations, respectively. In contrast, low flows have decreased in July and August at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Jul" & summary_table_direction_monthly$direction == "negative"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Aug" & summary_table_direction_monthly$direction == "negative"]` stations. Among them, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Jul" & summary_table_cat_monthly$category == "> 5% decrease"]` stations in July and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Low Flow" & summary_table_cat_monthly$Month == "Aug" & summary_table_cat_monthly$category == "> 5% decrease"]` in August show a decline at more than 5% per decade.

## Monthly peak flows

```{r fig.width = 10, fig.height=8, echo = FALSE}
monthly_peak_flow_bar_plot
```

-   Changes in monthly peak flows also vary by month. Peak flows have increased in January and April at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Apr" & summary_table_direction_monthly$direction == "positive"]` stations, respectively. In contrast, they have decreased in July and August at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Jul" & summary_table_direction_monthly$direction == "negative"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Aug" & summary_table_direction_monthly$direction == "negative"]` stations. Among them, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Jul" & summary_table_cat_monthly$category == "> 5% decrease"]` stations in July and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Peak Flow" & summary_table_cat_monthly$Month == "Aug" & summary_table_cat_monthly$category == "> 5% decrease"]` in August show a decline at more than 5% per decade.

\newpage

## Useful Links

## Data

The data sets are all sourced from the B.C. Data Catalogue, distributed
under the [Open Government Licence - British
Columbia](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61).

<!-- -   [Indicator Summary Data: Long-term Trends in Timing and Volume of
    B.C.'s River
    Flow](https://catalogue.data.gov.bc.ca/dataset/indicator-summary-data-change-in-timing-and-volume-of-river-flow-in-bc-1912-2012-) -->

-   River Flow Stations and daily flow data accessed with ['tidyhydat'](https://cran.r-project.org/web/packages/tidyhydat/index.html) R
    Package

<!-- Here Jon needs to upload watershed and basin layers (or at least point in right direction) -->
<!-- -   [Hydrologic -->
<!--     Zones](https://catalogue.data.gov.bc.ca/dataset/bc-hydrologic-zones). -->

------------------------------------------------------------------------

Published and Available Online at Environmental Reporting BC
(`r updateDate`):
<https://www.env.gov.bc.ca/soe/indicators/climate-change/rivers.html>

Email correspondence to:
[envreportbc\@gov.bc.ca](mailto:envreportbc@gov.bc.ca){.email}

*Suggested Citation*:\
Environmental Reporting BC. r format(Sys.Date(), "%Y"). Long-Term Trends in River Flows in
B.C. State of Environment Reporting, Ministry of Environment and Climate
Change Strategy, British Columbia, Canada.

## Methods

### Regime designation

The seasonal distribution of high flows and lows flow varied across the province. amounts and timing of precipitation, snowmelt, or both drove these variation. Their effects differed by station. So we used different flow metrics to represent hydrological changes. To assign flow metrics to each station, we determined seasonal streamflow regimes. We assessed these regimes using normalized monthly flows for each station^1^. We then applied a statistical clustering analysis to group the stations. This resulted in four high-level seasonal regimes.

### Filtering procedure

Monitoring varied across stations and rivers. To reduce errors in trend estimations, we applied a set of filtering procedures. These included:

-   **Missing annual data**: We removed station-years with more than 4% (~ 2 week) of missing data from the trend analysis. After this step, most stations had complete monthly data.

-   **Data gaps**: We removed all annual data preceding the gap of 5 or more years at each station. This step ensured continuous datasets across all stations for trend analyses.

-   **Recently active**: We included the stations that were active as recently as 2019 in the trend analysis.

-   **Minimum sample size**: We retained the stations with more than 10 years of monitoring data.

-   **Regulated Rivers**: We defined a river as a regulated if its flow regime changed, causing in a 10% increase or decrease in flow. We removed regulated stations from the analysis. But we kept stations with flow that we determined was close to natural.

-   **Upstream stations**: Many rivers include more than one monitoring station. In these cases, we kept only the downstream stations in the plots. The interactive map included an option to explore upstream locations.

-   **Recently installed stations**: The interactive map included an option to inspect stations installed after 1992. These stations had at least 10 years of data.

-   **Calendar year, water year, and low flow year**: Calendar year is the period between January 1st and December 31th. We defined water year as the period between October 1st of one year and September 30th of the next^2^. We defined low flow year as the period starting from April 1st to March 31st^3^.

### Data imputation

We used multivariate imputation by chained equations (MICE) to fill in missing values. We conducted MICE with predictive mean matching (PMM) to generate five imputed datasets. For each imputed dataset, we calculated six flow metrics. The volume metrics were mean annual flow, monthly averaged flow, low summer low and peak low. The timing metrics included date of freshet, and date of low flow. Finally, we pooled estimates from the imputed datasets using  Rubin’s rule. All imputation steps were conducted using mice R package. 
    
### Trend Analysis

We applied trend analysis to the pooled estimates at both annual and month scales. For the annual analysis, we calculated the mean value for each year across stations. For the monthly analysis, we calculated the mean value for each month over the time series. We used Kendall’s nonparametric test to assess trends using EnvStats R package. We estimated the trend magnitude using the Theil and Sen approach ^5,6^. 

We used p-value of 0.05 as the threshold of statistical significance. When p < 0.05, it indicated that there was < 5% likelihood of observing the trend by chance. When a trend was not significant (p > 0.05), it meant that it was not clear enough to be a trend. In such cases, we should interpret presence and magnitude with caution . Several factors can influence p-values. These included data variability, sample size, trend magnitude, and other factors. 

### References

- ^1^ Curran, J. H., & Biles, F. E. (2021). Identification of seasonal streamflow regimes and streamflow drivers for daily and peak flows in Alaska. Water Resources Research, 57. https://doi.org/10.1029/2020wr028425

-   ^2^[Description of Water Year (USGS)](https://water.usgs.gov/nwc/explain_data.html)

-   ^3^[Description of Low Flow Year (EPA)](https://www.epa.gov/hydrowq/definition-and-characteristics-low-flows)

-   ^4^ Theil, H. (1950). A Rank-Invariant Method of Linear and Polynomial Regression Analysis, I-III.Proc. Kon. Ned. Akad. v. Wetensch. A. 53, 386-392, 521-525, 1397-1412. https://link.springer.com/chapter/10.1007/978-94-011-2546-8_20

-   ^5^ Sen, P.K. (1968). Estimates of the Regression Coefficient Based on Kendall’s Tau. Journal of the American Statistical Association 63, 1379-1389. https://www.tandfonline.com/doi/abs/10.1080/01621459.1968.10480934

\newpage

## Appendix 1

### Changes in Volume and Timing by Watersheds

### Flow volume

### Average annual flow

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
average_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Average Annual Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

average_by_zone_direction = average_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
annual_bar_plot
```

-   For average annual flow, the vast majority of stations had non-significant trends. In the south-central and -west hydrologic zones, a number of stations had increasing annual flow trends including the Fraser Plateau (`r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Fraser Plateau" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Fraser Plateau"])` stations) and the west and central South Coast Mountains (`r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Western South Coast Mountains" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Western South Coast Mountains"])` and `r average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Central South Coast Mountains" & average_by_zone_direction$direction == "positive"]` of `r sum(average_by_zone_direction$n[average_by_zone_direction$Sub_Basin == "Central South Coast Mountains"])` stations respectively).

## Peak flow

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
peak_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Peak Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

peak_by_zone_direction = peak_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
peak_bar_plot
```

-   There were no clear geographic patterns in trends in peak flow. The only result to highlight is Western Vancouver Island, which has `r peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island" & peak_by_zone_direction$direction == "positive"]` of `r sum(peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations with increasing and `r peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island" & peak_by_zone_direction$direction == "negative"]` of `r sum(peak_by_zone_direction$n[peak_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations with decreasing annual peak flows.

## Low flow
```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
low_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Low Summer Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

low_by_zone_direction = low_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```


```{r fig.width = 10, fig.height=8, echo = FALSE}
low_bar_plot
```

-   Trends is summer low flow are most apparent on Vancouver Island. Western Vancouver Islands monitored rivers are almost all (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western Vancouver Island" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western Vancouver Island"])` stations) are decreasing at more than 5% volume per decade. Eastern Vancouver Island also has a large proportion of river low flows in decline (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Eastern Vancouver Island" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Eastern Vancouver Island"])` stations). Other hydrologic zones with decreasing summer low flow volumes to highlight include the Western South and Central Coast Mountains (`r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western South Coast Mountains" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Western South Coast Mountains"])` and `r low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Central Coast Mountains" & low_by_zone_direction$direction == "negative"]` of `r sum(low_by_zone_direction$n[low_by_zone_direction$Sub_Basin == "Central Coast Mountains"])` stations respectively)

# Flow timing

## Date of freshet

```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
freshet_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Date of Freshet") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

freshet_by_zone_direction = freshet_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```

```{r fig.width = 10, fig.height=8, echo = FALSE}
freshet_bar_plot
```

-   Timing of freshet has been getting earlier for stations across most hydrologic zones. North-eastern B.C. was the only region that was immune to this effect, apart from the Rocky Mountains which saw earlier freshet at several stations (`r freshet_by_zone_direction$n[freshet_by_zone_direction$Sub_Basin == "Northern Rocky Mountains" & freshet_by_zone_direction$direction == "negative"]` of `r sum(freshet_by_zone_direction$n[freshet_by_zone_direction$Sub_Basin == "Northern Rocky Mountains"])` stations). This pattern was also missing from Vancouver Island, although this is likely due to the majority of rivers not being snow and glacier-fed - the driver behind the freshet phenomena. 

## Start of low flow period
```{r echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}
date_low_by_zone_cat = mk_results_tbl %>%
  filter(metric == "Date of Low Summer Flow") %>%
   filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(Sub_Basin, category) %>%
  summarise(n = n())

date_low_by_zone_direction = date_low_by_zone_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(Sub_Basin, direction) %>%
  summarise(n = sum(n))
```

```{r fig.width = 10, fig.height=8, echo = FALSE}
date_low_bar_plot
```

Trends in the timing of the start of low flow period are far less clear than that of timing of freshet. Trends indicating earlier timing of low flows were present in some hydrologic zones, most prominent in the Southern Quesnel Highland (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Southern Quesnel Highland" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Southern Quesnel Highland"])` stations), the Fraser Plateau (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Fraser Plateau" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Fraser Plateau"])` stations) and Nechako Plateau (`r date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Nechako Plateau" & date_low_by_zone_direction$direction == "negative"]` of `r sum(date_low_by_zone_direction$n[date_low_by_zone_direction$Sub_Basin == "Nechako Plateau"])` stations).


## Appendix 2

```{r station_plots, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, results = 'asis', fig.height = 7.5, dev = "pdf"}

# individual well plots
# load('../groundwater-levels-indicator/tmp/well_plots.RData')

knitr_latex_char <- function(x) {
  y <- gsub("\\\\", "\\\\textbackslash{}", x) # backslash has to be first!
  y <- gsub("([#$%&_])", "\\\\\\1", y) # Doesn't deal with { or } because of function{}
  y <- gsub("\\^", "\\\\textasciicircum{}", y)
  y <- gsub("~", "\\\\textasciitilde{}", y)
  return(y)
}

# theme_set(base_theme)

  for (station in stations_sf_filtered$STATION_NUMBER) {
    
    stationname <- stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(STATION_NAME)
    
    regimetype = stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(Regime)
    
    reg = stations_sf_filtered %>%
      filter(STATION_NUMBER == station) %>%
      pull(Sub_Basin)
    
    cat(paste0('\\subsubsection*{',reg,": ",
               knitr_latex_char(stationname),
               '}'))
    
    cat(paste0('\\subsubsection*{',"Regime Type: ",
               knitr_latex_char(regimetype),
               '}'))

    mapplot <- readPNG(here(paste0("print_ver/out/figs/maps/",stationname,".png")))
    mapplot = ggdraw() +
      draw_image(mapplot)
    
    hydroplot = readPNG(here(paste0("print_ver/out/figs/hgraphs/",station,".png")))  
    hydroplot = ggdraw() +
      draw_image(hydroplot)
    
    volumeplot = readPNG(here(paste0("print_ver/out/figs/volume_plots/",station,".png")))
    volumeplot = ggdraw() +
      draw_image(volumeplot)
    
    timingplot = readPNG(here(paste0("print_ver/out/figs/timing_plots/",station,".png")))  
    timingplot = ggdraw() +
      draw_image(timingplot)    
    
    grid.arrange(mapplot+ theme(plot.margin = unit(c(-1,0.5,-1,0.01),"cm")),
                 hydroplot+ theme(plot.margin = unit(c(-1,0.01,-1,0.01),"cm")),
                 volumeplot + theme(plot.margin = unit(c(0.5,0.5,-1,0.01),"cm")),
                 timingplot + theme(plot.margin = unit(c(0.5,0.5,0,0.01),"cm")),
                 ncol = 2)
    # 
    cat('\\newpage ')
  }
```
