---
title: Long-Term Trends in River Flows in B.C.
topic: "water"
format:
     pdf: 
       toc: true
       toc-location: left
       template: "template.tex"
       keep-tex: FALSE
       latex_engine: xelatex
       gphxpath: "img/"
       mainfont: "Palatino Linotype"
       sansfont: "Calibri"
     html:
       theme: none
       embed-resources: true
       encoding: 'UTF-8'
       include-in-header: 
          - file: "flow_files/head.html"
          - file: "flow_files/header.html"
          - file: "flow_files/header-water.html"
       #include-before-body:
       include-after-body:
          - file: "flow_files/footer.html"
       editor: visual
---

```{=html}
<!--
Copyright 2025 Province of British Columbia

Licensed under the Apache License, Version 2.0 (the "License");
you may not use this file except in compliance with the License.
You may obtain a copy of the License at

http://www.apache.org/licenses/LICENSE-2.0

Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
See the License for the specific language governing permissions and limitations under the License.
-->
```

\pagebreak

```{r setup, echo = FALSE, warning = FALSE, message = FALSE, cache = FALSE, results = 'hide'}

library(dplyr)
library(ggplot2)
library(grid)
library(RColorBrewer)
library(scales)
library(bcgroundwater)
library(envreportutils)
library(gridExtra)
library(xtable)
library(lubridate)
library(forcats)
library(here)
library(magick)
library(png)
library(cowplot)
library(here)
library(sf)
library(stringr)

knitr::opts_chunk$set(cache = FALSE)

theme_set(theme_classic() +
            theme(text = element_text(colour = "black"),
                  axis.line = element_blank(),
                  axis.ticks = element_blank(),
                  panel.grid.major = element_line(colour = "grey85", size = 0.5,
                                                linetype = 1),
                  panel.grid.minor = element_line(colour = "grey90", size = 0.5,
                                                linetype = 1),
                  panel.grid.major.x = element_blank(),
                  panel.spacing = unit(0.6, "lines"),
                  plot.title = element_text(vjust = 2, hjust = 0.5),
                  axis.title = element_text(vjust = 0.1),
                  legend.position = "bottom", legend.title = element_blank(),
                  legend.text = element_text(size = 12),
                  axis.text.x = element_blank(),
                  strip.background = element_blank()))

label.colour <- "black" 
colour.scale <- brewer.pal(3,"Blues")

updateDate <- "February 2024"

# Load datasets
annual_flow_dat = readRDS(here("app/www","annual_flow_dat.rds"))
stations_sf = st_read(here("app/www","stations.gpkg")) 
mk_results_tbl = readRDS(here("print_ver/out/mk_results_tbl.rds"))
mk_results_monthly_tbl = readRDS(here("print_ver/out/mk_results_monthly_tbl.rds"))

stations_sf_filtered = stations_sf %>%
  filter(keep == 1)

# summary table of number of stations in each category
summary_table_cat = mk_results_tbl %>%
  filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(metric, Regime, category) %>%
  summarise(n = n())

summary_table_direction = summary_table_cat %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(metric, Regime,  direction) %>%
  summarise(n = sum(n))

#Subset Average
average_cat = summary_table_cat %>%
  filter(metric == "Average Annual Flow")

average_direction = summary_table_direction %>%
  filter(metric == "Average Annual Flow")

#Subset Peak Flow
peak_cat = summary_table_cat %>%
  filter(metric == "Peak Flow")

peak_direction = summary_table_direction %>%
  filter(metric == "Peak Flow")

#Subset Low Flow
Low_cat = summary_table_cat %>%
  filter(metric == "Low Summer Flow")

Low_direction = summary_table_direction %>%
  filter(metric == "Low Summer Flow")

#Subset Freshet
freshet_cat = summary_table_cat %>%
  filter(metric == "Date of Freshet*")

freshet_direction = summary_table_direction %>%
  filter(metric == "Date of Freshet*")

#Subset Freshet
low_date_cat = summary_table_cat %>%
  filter(metric == "Start of Low Flow Period")

low_date_direction = summary_table_direction %>%
  filter(metric == "Start of Low Flow Period")

# Results by regime
annual_regime = 

# Monthly results

# Subset median flow by month
summary_table_cat_monthly = mk_results_monthly_tbl %>%
  filter(STATION_NUMBER %in% stations_sf_filtered$STATION_NUMBER) %>%
    mutate(category = case_when(significant == 0.1 ~ "No significant trend",
                                     .default = magnitude_fixed)) %>%
  group_by(metric, Month, category) %>%
  summarise(n = n())

summary_table_direction_monthly = summary_table_cat_monthly %>%
  mutate(direction = case_when(str_detect(category,"increase") | str_detect(category,"later")  ~ "positive",
                               str_detect(category, "decrease") | str_detect(category, "earlier") ~ "negative",
                               .default = "no change")) %>%
  group_by(metric, Month, direction) %>%
  summarise(n = sum(n))

# Load figures
load(here("print_ver/out","figures.Rdata"))

# Load map png (perhaps change this to a nicer looking map?)
leaflet_map = image_read(here("print_ver/out/figs","static_leaflet.png"))

is_html <- knitr::is_html_output()
is_pdf <- knitr::is_latex_output()
```

-   **Rivers are one of the most important water sources for people living in British Columbia.** Measuring the flow of rivers across B.C. is essential for effective management and sustainable use.

-   **Changes in the timing and volume of river flow can affect both human and natural ecosystems.** These changes can affect our ability to predict and manage seasonal water resources. They may disrupt natural systems that rely on typical flow patterns.

-   **Low river flows in summer may reduce the amount of water available for use. They include agriculture, hydroelectric power generation, industry and communities.** Low flows may impede salmon migration and often link to warmer water temperatures. Reduced river levels can decrease water quality, threatening the health of aquatic ecosystems.

-   **Extreme high river flows can pose risks to human life and may lead to economic losses.** Flood events are often linked to rainfall and snowmelt.

-   **Long-term trends in the river flow volume vary by flow metric and regime type.** Trends in peak flow volume are stable across the province. But annual flows are increasing for many snow-dominated rivers. Summer low flow volumes are decreasing, particularly in rivers with rain-dominated regimes.

-   **Freshet is occurring earlier for a large number of rivers across the province.** This change has a less pronounced effect on the start of low flow period. But several snow-dominated rivers are still showing an earlier start to this period.

-   **There has been a shift in seasonal timing of flow events, as seen in monthly flow trends.** Average and summer peak flows are increasing during the winter and spring (November - May). They are declining in the summer (July - September).

<!-- -   **Trends in timing and volume of river flow vary geographically.** Annual average flows increased in the south-west of the province, while there was a strong trend for lower river flows on Vancouver Island. Timing of freshet was consistently earlier across all regions apart from the North-East.  -->

\newpage

::: {.content-visible when-format="html"}
## On this page

-   <a href="#river-flows-in-british-columbia" id="toc-river-flows-in-british-columbia"><u>River flows in British Columbia</u></a>

-   <a href="#description-of-river-flow-metrics" id="toc-description-of-river-flow-metrics"><u>Description of river flow metrics</u></a>

-   <a href="#river-stations" id="toc-river-stations"><u>River Stations</u></a>

-   <a href="#trend-results" id="toc-trend-results"><u>Trend Results</u></a>

-   <a href="#data" id="toc-data"><u>Data, Methods, and References</u></a>
:::

To help understand the trends measured in rivers across the province, there are several key foundation

Flow Regimes across British Columbia

### Types of flow regimes

Flow volume and timing of specific flow events vary based on the source of the river water. There are many types of river flow regimes^1^. This analysis divides the regimes into four categories:

-   **Snow-Dominated - Early Peak**: Snowmelt in the spring and early summer is a major source of flows. This causes a spring freshet and lower volumes in summer.

-   **Snow-Dominated - Late Peak**: Flows are also driven by snowmelt. But, due to higher latitude and elevations at the river source, snowmelt occurs later in the year. This leads to a delayed freshet and later summer low flows.

-   **Rain-Dominated**: Precipitation is the main source of flows. In B.C., these rivers are in the coastal and island regions. Such rivers typically flow high in the winter and low in the summer.

-   **Mixed Regime**: Flows come from both snowmelt and precipitation. This leads to a freshet event and high river flows in the winter.

## Description of river flow metrics

The metrics calculated at each station aim to quantify two aspects of river flow changes: 1) the volume of river flow and 2) the timing of key river flow events.

### Volume metrics

#### Average annual flow

Average annual flow (meters^3^/s) is the daily flow averaged for each station and water year^2^ (October – October).

#### Peak annual flow

Volume of peak flow (meters^3^/s) is the highest flow identified from the rolling averages with a 3-day window. This metric represents the peak flow for each station for each water year^2^ (October - October).

#### Low summer flow

Volume of low flows (meters^3^/s) represents the lowest flow for each low-flow year^3^ at each station. The metric uses the smallest rolling average with a 7-day window across the summer months (May - October).

### Timing metrics

#### Date of Freshet

We define the freshet as the seasonal flood of a river, often due to snow melt. The freshet date is defined as the day when cumulative river discharge reaches 50% of the total for a specified period. The effects of snowmelt on river flow vary by regime. So the method of calculating freshet timing differs among regimes. For snow-dominated regimes, we use water year to account for early peak. For late-peak regimes, the calendar year is used to reflect the delayed freshet. For mixed regimes, we exclude the rain-dominated winter months (October - February). Instead, freshet timing is estimated based on discharge calculated between March and September. For rain-dominated stations, we do not calculate freshet timing, as freshet does not take place.

#### Start of low flow period

We use the long-term Mean Annual Discharge (MAD) to estimate the beginning of low summer flows. The start of the low-flow period is defined as the first day when discharge drops to or below 50% of MAD after peak flow. For most rivers, a 50% MAD threshold is used, but for rain-dominated rivers, a lower threshold of 20% MAD is applied.

\newpage

## River stations

This analysis considers `r nrow(stations_sf_filtered)` stations across the Province, or `r nrow(stations_sf)` stations when including upstream stations. These stations have tracked river flow for `r min(stations_sf_filtered$N_years)` to `r max(stations_sf_filtered$N_years)` years between `r min(stations_sf_filtered$Min_Year)` and `r max(stations_sf_filtered$Max_Year)`.

The map below shows all `r nrow(stations_sf_filtered)` downstream river stations. The color of the dots represents the flow regime of the rivers where the stations are located. Black lines show the boundaries of the six major river basins. They, from top left to bottom right, are Northwest, Liard, Peace, Coastal, Fraser, and Columbia. Grey lines outline 23 sub-basins.

```{r include = is_html, eval=is_html, overview_map, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.align = 'left', fig.width = 28, fig.asp=0.75}
plot(leaflet_map)
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height = 6, echo = FALSE, fig.alt="Leaflet map"}
plot(leaflet_map)
```

## Trend results

::: {.content-visible when-format="html"}
### Exploring water flow trends

The interactive map below offers several options to explore.

-   To begin, go to 'Trend options' tab in the bottom left corner. You can select a basin of interest from the drop-down menu or by clicking it on the map. Choose annual trend or explore the trend for a specific month by selecting ‘Monthly’. Next, navigate to the 'Trend to Display' option to select a flow metric of interest from the dropdown menu. You can choose a timespan: 'Three decades (1992–2022)' or 'All available years'. Check the two boxes to explore recently installed or upstream stations. This helps focus on newer stations or compare trends between upstream and downstream stations.

-   After selecting the 'Trend Options', click on a station of interest on the map. This will display the flow metric plot in the bottom left corner.

-   To explore hydrograph information for the selected station, go to the ‘Station Hydrograph’ tab.

-   To reset the view, select “All” or click on the B.C. button in the top left corner to return to all regions.

```{=html}
<iframe style="width: 100%; height: 1000px; border: none" frameborder="0" data-external="1" src="https://bcgov-env.shinyapps.io/flow-indicator-2025/" id="Interactive visualization of Long-Term Trends in River Flows in B.C."></iframe>
```
:::

### Volume metrics

-   Of the `r nrow(stations_sf_filtered)` monitored stations, `r sum(average_direction$n[average_direction$direction == "positive"])` showed an increase in average annual flows. Among them, `r sum(average_cat$n[average_cat$category == "> 5% increase"])` rose by more than 5% per decade. In contrast, `r sum(average_direction$n[average_direction$direction == "negative"])` stations had decreasing average annual flows. The remaining `r sum(average_direction$n[average_direction$direction == "no change"])` stations had no significant trend.

-   Annual peak flows increased at `r sum(peak_direction$n[peak_direction$direction == "positive"])` stations and decreased at `r sum(peak_direction$n[peak_direction$direction == "negative"])` stations. The vast majority (`r sum(peak_direction$n[peak_direction$direction == "no change"])` stations) showed no significant change.

-   For low summer flow volume, `r sum(Low_direction$n[Low_direction$direction == "negative"])` showed a decline. `r sum(Low_cat$n[Low_cat$category == "> 5% decrease"])` of them decreased at a rate greater than 5% per decade.

```{r include = is_html, eval=is_html, fig.width = 10, fig.height = 8, echo = FALSE, fig.alt="MK_all"}
mk_all
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height = 6, fig.alt="MK_all", echo = FALSE}
mk_all
```

-   Date of Freshet plot uses a reduced number of stations. This is because the analysis does not include rain-dominated rivers.

### Timing metrics

-   Freshet had been occurring earlier for many locations. Among them, `r sum(freshet_direction$n[freshet_direction$direction == "negative"])` showed an earlier trend. Of these, `r sum(freshet_cat$n[freshet_cat$category == "> 2 days earlier"])` stations were occurring 2 days earlier per decade. `r sum(freshet_direction$n[freshet_direction$direction == "no change"] + freshet_cat$n[freshet_cat$category == "< 1 days change"])` stations showed little change, with less than 1 day change per decade, or no significant trend. No stations showed a late freshet.

-   Most locations did not show significant trends in the start date of the low flow period. Among those that did, `r sum(low_date_direction$n[low_date_direction$direction == "negative"])` stations had earlier low flows, while only `r sum(low_date_direction$n[low_date_direction$direction == "positive"])` stations had later low flows.

### Changes in flow volume by month

### Average monthly flows

```{r include = is_html, eval=is_html, fig.width = 10, fig.height=8, echo = FALSE, fig.alt="Monthly averaged bar plot"}
monthly_average_bar_plot
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height=6, fig.alt="Monthly averaged bar plot", echo = FALSE}
monthly_average_bar_plot + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(nrow = 2))
```

-   Changes in average river flow across years varied by month. Many rivers showed increasing average flows in winter and spring months (December - May). For example, January flows were increasing at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Average" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` stations. In April and May, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Apr" & summary_table_cat_monthly$category == "> 5% increase"]` and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "May" & summary_table_cat_monthly$category == "> 5% increase"]` stations showed a more than 5% increase per decade.

### Monthly low flows

```{r include = is_html, eval=is_html, fig.width = 10, fig.height=8, echo = FALSE, fig.alt="Monthly low flow bar plot"}
monthly_low_flow_bar_plot
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height=6, fig.alt="Monthly low flow bar plot", echo = FALSE}
monthly_low_flow_bar_plot +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(nrow = 2))
```

-   Low flows had increased in January and April at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Apr" & summary_table_direction_monthly$direction == "positive"]` stations, respectively. In contrast, low flows had decreased in July and August at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Jul" & summary_table_direction_monthly$direction == "negative"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Low Flow" & summary_table_direction_monthly$Month == "Aug" & summary_table_direction_monthly$direction == "negative"]` stations. Among them, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Jul" & summary_table_cat_monthly$category == "> 5% decrease"]` stations in July and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Low Flow" & summary_table_cat_monthly$Month == "Aug" & summary_table_cat_monthly$category == "> 5% decrease"]` in August showed a decline of more than 5% per decade.

### Monthly peak flows

```{r include = is_html, eval=is_html, fig.width = 10, fig.height=8, echo = FALSE, fig.alt="Monthly peak flow bar plot"}
monthly_peak_flow_bar_plot
```

```{r include = is_pdf, eval=is_pdf, fig.width = 7, fig.height=6, fig.alt="Monthly peak flow bar plot", echo = FALSE}
monthly_peak_flow_bar_plot +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  guides(fill = guide_legend(nrow = 2))
```

-   Changes in monthly peak flows also varied by month. Peak flows had increased in January and April at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Jan" & summary_table_direction_monthly$direction == "positive"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Apr" & summary_table_direction_monthly$direction == "positive"]` stations, respectively. In contrast, they had decreased in July and August at `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Jul" & summary_table_direction_monthly$direction == "negative"]` and `r summary_table_direction_monthly$n[summary_table_direction_monthly$metric == "Peak Flow" & summary_table_direction_monthly$Month == "Aug" & summary_table_direction_monthly$direction == "negative"]` stations. Among them, `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Average" & summary_table_cat_monthly$Month == "Jul" & summary_table_cat_monthly$category == "> 5% decrease"]` stations in July and `r summary_table_cat_monthly$n[summary_table_cat_monthly$metric == "Peak Flow" & summary_table_cat_monthly$Month == "Aug" & summary_table_cat_monthly$category == "> 5% decrease"]` in August showed a decline of more than 5% per decade.

\newpage

::: {style="background-color: rgba(0, 51, 102, 0.80)"}
## [Methods]{style="color:#fff;"}
:::

::: {.content-visible when-format="pdf"}
### Regime designation

The seasonal distribution of high flows and low flows varied across the province. Amounts and timing of precipitation, snowmelt, or both drove this variation. Their effects differed by station. So we used different flow metrics to represent hydrological changes. To assign flow metrics to each station, we determined seasonal streamflow regimes. We assessed these regimes using normalized monthly flows for each station^1^. We then applied a statistical clustering analysis to group the stations. This resulted in four high-level seasonal regimes.

### Filtering procedure

Monitoring varied across stations and rivers. To reduce errors in trend estimations, we applied a set of filtering procedures. These included:

-   **Missing annual data**: We removed station-years with more than 4% (\~ 2 weeks) of missing data from the trend analysis. After this step, most stations had complete monthly data.

-   **Data gaps**: We removed all annual data preceding the gap of 5 or more years at each station. This step ensured continuous datasets across all stations for trend analyses.

-   **Recently active**: We included the stations that were active as recently as 2019 in the trend analysis.

-   **Minimum sample size**: We retained the stations with more than 10 years of monitoring data.

-   **Regulated Rivers**: We defined a river as a regulated if its flow regime changed, causing a 10% increase or decrease in flow. We removed regulated stations from the analysis. But we kept stations with the flow that we determined was close to natural.

-   **Upstream stations**: Many rivers included more than one monitoring station. In these cases, we kept only the downstream stations in the plots. The interactive map included an option to explore upstream locations.

-   **Recently installed stations**: The interactive map included an option to inspect stations installed after 1992. These stations had at least 10 years of data.

-   **Calendar year, water year, and low flow year**: Calendar year is the period between January 1st and December 31st. We defined water year as the period between October 1st of one year and September 30th of the next^2^. We defined low flow year as the period starting from April 1st to March 31st^3^.

### Data imputation

We used multivariate imputation by chained equations (MICE) to fill in missing values. We conducted MICE with predictive mean matching (PMM) to generate five imputed datasets. For each imputed dataset, we calculated six flow metrics. The volume metrics were mean annual flow, monthly averaged flow, low summer low, and peak low. The timing metrics included date of freshet and date of low flow. Finally, we pooled estimates from the imputed datasets using Rubin’s rule. All imputation steps were conducted using mice R package.

### Trend analysis

We applied trend analysis to the pooled estimates at both annual and monthly scales. For the annual analysis, we calculated the mean value for each year across stations. For the monthly analysis, we calculated the mean value for each month over the time series. We used Kendall’s nonparametric test to assess trends using EnvStats R package. We estimated the trend magnitude using the Theil and Sen approach ^4,5^.

We used the p-value of 0.05 as the threshold of statistical significance. When p \< 0.05, it indicated that there was a \< 5% likelihood of observing the trend by chance. When a trend was not significant (p \> 0.05), it meant that it was not clear enough to be a trend. In such cases, we should interpret its presence and magnitude with caution. Several factors can influence p-values. These included data variability, sample size, trend magnitude, and other factors.
:::

::: {.content-visible when-format="html"}
-   Detailed methods used to develop this indicator, along with relevant results (e.g., volume and timing of flow by river regime), are available in the [printable version of this indicator (PDF)](http://wwwd.env.gov.bc.ca/soe/indicators/water/print_ver/envreportbc_FI_April2025.pdf)

-   Current trend analysis methods differ from those used in the previous version of this flow indicator, so the results are not directly comparable.
:::

::: {style="background-color: rgba(0, 51, 102, 0.80)"}
## [References]{style="color:#fff;"}
:::

-   ^1^ Curran, J. H., & Biles, F. E. (2021). Identification of seasonal streamflow regimes and streamflow drivers for daily and peak flows in Alaska. Water Resources Research, 57. https://doi.org/10.1029/2020wr028425

-   ^2^[Description of Water Year (USGS)](https://water.usgs.gov/nwc/explain_data.html)

-   ^3^[Description of Low Flow Year (EPA)](https://www.epa.gov/hydrowq/definition-and-characteristics-low-flows)

-   ^4^ Theil, H. (1950). A Rank-Invariant Method of Linear and Polynomial Regression Analysis, I-III.Proc. Kon. Ned. Akad. v. Wetensch. A. 53, 386-392, 521-525, 1397-1412. https://link.springer.com/chapter/10.1007/978-94-011-2546-8_20

-   ^5^ Sen, P.K. (1968). Estimates of the Regression Coefficient Based on Kendall’s Tau. Journal of the American Statistical Association 63, 1379-1389. https://www.tandfonline.com/doi/abs/10.1080/01621459.1968.10480934

::: {style="background-color: rgba(0, 51, 102, 0.80)"}
## [Data]{style="color:#fff;"}
:::

The data sets are all sourced from the B.C. Data Catalogue, distributed under the [Open Government Licence - British Columbia](https://www2.gov.bc.ca/gov/content?id=A519A56BC2BF44E4A008B33FCF527F61).

```{=html}
<!-- -   [Indicator Summary Data: Long-term Trends in Timing and Volume of
    B.C.'s River
    Flow](https://catalogue.data.gov.bc.ca/dataset/indicator-summary-data-change-in-timing-and-volume-of-river-flow-in-bc-1912-2012-) -->
```

-   River Flow Stations and daily flow data accessed with ['tidyhydat'](https://cran.r-project.org/web/packages/tidyhydat/index.html) R Package

<!-- Here Jon needs to upload watershed and basin layers (or at least point in right direction) -->

<!-- -   [Hydrologic -->

<!--     Zones](https://catalogue.data.gov.bc.ca/dataset/bc-hydrologic-zones). -->

------------------------------------------------------------------------

Published and Available Online at Environmental Reporting BC (`r updateDate`)

::: {.content-visible when-format="html"}
[Download a printable version of this indicator (PDF)](http://wwwd.env.gov.bc.ca/soe/indicators/water/print_ver/envreportbc_FI_April2025.pdf)
:::

Email correspondence to: [envreportbc\@gov.bc.ca](mailto:envreportbc@gov.bc.ca){.email}

*Suggested Citation*:\
Environmental Reporting BC. `r format(Sys.Date(), "%Y")`. Long-Term Trends in River Flows in B.C. State of Environment Reporting, Ministry of Environment and Climate Change Strategy, British Columbia, Canada.

\newpage

```{r child=if (is_pdf) here("print_ver", "flow-indicator-appendix.qmd")}
```
